---
title: "`pig_latinator` function"
author: Javier Mtz.-Rdz.
output: github_document
---

# Setup

For this assignment, we will need to load the following packages.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(testthat)
```

# Creating a Function for Pig Latin Transformation

In this section, I developed a function to convert words or sentences into a customized Pig Latin format.

*Component Implementation*. The function implements sthe rules present on [Wikipedia](https://en.wikipedia.org/wiki/Pig_Latin):

-   For words starting with vowels, it moves initial vowel(s) along with the first consonant or consonant cluster to the end.
-   For words starting with consonants, it moves the initial consonant or consonant cluster to the end.

*Addition Component*. The additional components include:

-   It retains punctuation at the start and end of words.
-   Handles uppercase logic, converting words to uppercase or sentence case as needed.
-   Allows the choice between returning a vector of words or a concatenated string through the concat argument.

```{r}
#' Pig Latinator Function
#'
#' The `pig_latinator` function converts words or sentences to a form of Pig Latin.
#' It retains punctuation and uppercase logic (single-letter words are always 
#' treated as capitalized words).
#' 
#' @param text A character vector containing the text to convert to Pig Latin.
#' @param addition_rule A character specifying the addition rule for Pig Latin. 
#' Default is "ay".
#' @param no_cons A character specifying the addition in case of no consonants 
#' (it should be a consonant). Default is "w".
#' @param concat Logical, indicating whether to concatenate the vector of words.
#' If `TRUE`, the function returns a single string. If `FALSE`, it returns a vector.
#' Default is `TRUE`.
#' 
#' @return A vector of words converted to Pig Latin, or a single string 
#' if `concat` is `TRUE`.
#' 
#' @seealso [Pig Latin on Wikipedia](https://en.wikipedia.org/wiki/Pig_Latin)
#' @export
#' @examples
#'# Basic example
#'pig_latinator("Hello world")
#'[1] "Ellohay orldway"
#'
#'# Handling punctuation
#'pig_latinator("This is a test. Does it work?")
#'[1] "Isthay isay away esttay. Oesday itay orkway?"
#'
#'# Managing uppercase and lowercase.
#'pig_latinator("Sentence with UPPER and lower case.")
#'[1] "Entencesay ithway ERUPPAY anday owerlay asecay."
#'
#'# Different arguments
#'# Managing uppercase and lowercase.
#'pig_latinator(c("This", "is a Pig Latin", "vector", " of words"), 
#'              addition_rule = "ey",
#'              no_cons = "x",
#'              concat = F)
#'[1] "Isthey"   "isey"     "axey"     "Igpey"    "Atinley"  "ectorvey" "EY"
#'[8] "ofey"     "ordswey" 
pig_latinator <- function(text,
                          addition_rule = "ay", 
                          no_cons = "w",
                          concat = TRUE) {
  # Check if input is a character vector
  if (!is.character(text)) {
    stop("Input 'text' must be a character vector.")
  }
  
  # Check if addition_rule is character
  if (!is.character(addition_rule)) {
    stop("Input 'addition_rule' must be a character.")
  }
  
  # Check if no_cons is character
  if (!is.character(no_cons)) {
    stop("Input 'no_cons' must be a character.")
  }
  
  # Check if no_cons is not a consonant
  if (!stringr::str_detect(no_cons, "[b-df-hj-np-tv-z]|[B-DF-HJ-NP-TV-Z]")) {
    stop("Input 'no_cons' must be a consonant.")
  }
  
  # Additions as lowercases
  addition_rule <- stringr::str_to_lower(addition_rule)
  no_cons <- stringr::str_to_lower(no_cons)
  
  # Split and unlist text (in case that more than a word is provided)
  word <- unlist(stringr::str_split(text, " "))
  
  # Vectorized process for each word
  word <- purrr::map_vec(
    word,
    function(w) 
    {
      # Remove punctuation at the end and start
      ent_punct <- stringr::str_extract(w, "[:punct:]$")
      strt_punct <- stringr::str_extract(w, "^[:punct:]")
      w <- stringr::str_remove_all(w, "^[:punct:]|[:punct:]$")
      
      # Detect uppercase logic
      cnt_lttr <- stringr::str_count(w, "[:alpha:]")
      upp_lttr <- stringr::str_count(w, "[:upper:]")
      upp_first <- stringr::str_count(w, "^[:upper:]")
      case_upp <- dplyr::case_when(cnt_lttr == 1 & cnt_lttr == upp_lttr ~ "sentence",
                                    cnt_lttr == upp_lttr ~ "uppercase",
                                    upp_first == 1 ~ "sentence",
                                    TRUE ~ "other")
      w <- stringr::str_to_lower(w)
      
      # Convert words beginning with vowels to Pig Latin
      if (stringr::str_detect(w, "^[aeiou]")) {
        ext <- stringr::str_extract(w, "^[aeiou]+[b-df-hj-np-tv-z]*")
        if (!stringr::str_detect(ext, "[b-df-hj-np-tv-z]$")){
          ext <- stringr::str_c(ext, no_cons)
        }
        w <- stringr::str_c(stringr::str_remove(w, "^[aeiou]+[b-df-hj-np-tv-z]*"),
                            ifelse(is.na(ext), "", ext))
      } 
      
      # Convert words beginning with consonants to Pig Latin
      if (stringr::str_detect(w, "^[b-df-hj-np-tv-z]")){
        ext <- stringr::str_extract(w, "^[b-df-hj-np-tv-z]+")
        w <- stringr::str_c(stringr::str_remove(w, "^[b-df-hj-np-tv-z]+"),
                            ifelse(is.na(ext), "", ext))
      }
      
      # Add addition_rule
      w <- stringr::str_c(w, addition_rule)
      
      # Add corresponding uppercases
      w <- dplyr::case_when(case_upp == "uppercase" ~ stringr::str_to_upper(w),
                            case_upp == "sentence" ~ stringr::str_to_sentence(w),
                            TRUE ~ w)
      
      # Add previously removed punctuation
      w <- stringr::str_c(ifelse(is.na(strt_punct), "", strt_punct),
                          w,
                          ifelse(is.na(ent_punct), "", ent_punct))
      
      return(w)
    })
  
  # Concatenate words in case of being true
  if (concat) word <- stringr::str_c(word, collapse = " ")
  
  return(word)
}

```

# Examples

The following examples show different uses of the `pig_latinator` function.

### Example 1: Translating "The Divine Comedy" to Pig Latin
In this example, we translate the beginning of ["The Divine Comedy" by Dante Alighieri from the Project Gutenberg](https://dev.gutenberg.org/ebooks/8800) into Pig Latin.

```{r}
# Download the text of "The Divine Comedy" by Dante Alighieri
book <- read.table("https://dev.gutenberg.org/cache/epub/8800/pg8800.txt",
                   sep =  "\t",
                   skip = 168, 
                   nrows = 17) %>% 
  rename(txt = 1) 

# Translate to Pig Latin 
translated_text <- pig_latinator(book$txt)

translated_text
```

### Example 2: Translating tidytext stopwords to Pig Latin
In this example, we translate [tidytext stopwords](https://github.com/juliasilge/tidytext) into Pig Latin. Unlike the previous example, we aim to obtain a vector of words.

```{r}
# Get tidytext stopwords
tdy_txt_sw <- tidytext::stop_words

# Display the first few stopwords
head(tdy_txt_sw)

# Translate to Pig Latin 
translated_stopwords <- pig_latinator(tdy_txt_sw$word,
                                      concat = FALSE)
head(translated_stopwords)
```

### Example 3: Translating "The Divine Comedy" with modified rules

Here, we translate the initial text of "The Divine Comedy" to Pig Latin using different rules for added letters.

```{r}
# Translate to Pig Latin with altered rules
modified_pig_latin <- pig_latinator(book$txt,
                                    addition_rule = "ei",
                                    no_cons = "x")
head(modified_pig_latin)
```


# Test `pig_latinator`

In this section, we teste the `pig_latinator` function.

```{r}

# Test 1: Check errors
test_that("Test errors", {
  input_text <- "This is a test"
  expect_error(pig_latinator(342),
               "Input 'text' must be a character vector.")
  expect_error(pig_latinator(input_text,
                             342),
               "Input 'addition_rule' must be a character.")
  expect_error(pig_latinator(input_text,
                             no_cons = 342),
               "Input 'no_cons' must be a character.")
  expect_error(pig_latinator(input_text,
                             no_cons = "a"),
               "Input 'no_cons' must be a consonant.")
})


# Test 2: Check if function retains uppercase logic and diferent rules
test_that("Uppercase logic is and punctuation retained", {
  input_text <- "HELLO WORLD! This is a test. I eat late."
  output1 <- pig_latinator(input_text)
  output2 <- pig_latinator(input_text,
                           addition_rule = "ei",
                           no_cons = "x")
  expect_equal(output1, 
               "ELLOHAY ORLDWAY! Isthay isay away esttay. Iway eatay atelay.")
  expect_equal(output2,
               "ELLOHEI ORLDWEI! Isthei isei axei esttei. Ixei eatei atelei.")
})

# Test 3: Check if function retains a vector of words
test_that("Returns vector of words", {
  input_text <- "This is a test"
  output <- pig_latinator(input_text, concat = FALSE)
  expect_length(output, 4)
  expect_is(output, "character")
})


```
